Make any stupid plot you want

### 2021-03-01 draft/sections/intro.tex
rework of nu

\subsection{Contribution}
This work presents a mathematical model of the transformation from data to graphic representation and a proof of concept implementation. Specifically, this work contributes

\begin{enumerate}
    \item a functional oriented visualization tool architecture
    \item topology-preserving maps from data to graphic 
    \item monoidal action equivariant maps from component to visual variable 
    \item algebraic sum such that more complex visualizations can be built from simple ones 
    \item prototype built on Matplotlib's infrastructure
\end{enumerate}
In contrast to mathematical models of visualization that aim to evaluate visualization design, we propose a topological framework for building tools to build visualizations. We defer judgement of expressivity and effectiveness to developers building domain specific tools, but provide them the framework to do so. 
\end{document}

>>>>

\subsection{Contribution}
This work presents a mathematical model of the transformation from data to graphic representation and a proof of concept implementation. Specifically, the contributions of this work are

\begin{enumerate}
    \item a formal description of the topology preserving relationship between data and graphic via continuous maps
    \item a formal description of the property preservation from data component to visual representation as equivariant maps that carry a homomorphism of monoid actions
    \item abstraction of data as a fiber bundle with schema like fibers to encode components and topology 
    \item algebraic sum operator such that more complex visualizations can be built from simple ones 
    \item a functional oriented visualization tool architecture built on the mathematical model to demonstrate the utility of the model
    \item a prototype of the architecture built on Matplotlib's infrastructure to demonstrate the feasibility of the model
\end{enumerate}

In contrast to mathematical models of visualization that aim to evaluate visualization design, we propose a topological framework for building tools to build visualizations. We defer judgement of expressivity and effectiveness to developers building domain specific tools, but provide them the framework to do so. 
\end{document}

### 2021-03-01 draft/sections/math.tex
rework of nu

As discussed in the introduction, visualization is generally defined as structure preserving maps from data to graphic representation. In order to formalize this statement, we describe the connectivity of the records using topology and define the structure on the components in terms of the monoid actions on the component types. By formalizing structure in this way, we can evaluate the extent to which a visualization preserves the structure of the data it is representing and build structure preserving visualization tools. 
We introduce the notion of an artist $\mathscr{\vartist}$ as an equivariant map from data to graphic
\begin{equation}
    \label{eq:artist}
    \mathscr{\vartist}: \mathscr{\dtotal} \rightarrow \mathscr{\gtotal}
\end{equation}
that carries a homomorphism of monoid actions $\varphi: \monoid \rightarrow \monoid^{\prime}$ \cite{cegarraCohomologyMonoidsOperators2019}, which are discussed in detail in section~\ref{sec:data_monoid}. Given \monoid\ on data $\mathscr{\dtotal}$ and $\monoid^{\prime}$ on graphic $\mathscr{\gtotal}$, we propose that artists $\mathscr{\vartist}$ are symmetric 
\begin{equation}
\mathscr{\vartist}(m\cdot \delement) = \varphi(m)\cdot\mathscr{\vartist}(\delement) 
\end{equation}
such that applying a monoid action $m \in \monoid$ to the input $\delement \in \mathscr{\dtotal}$ to $\mathscr{\vartist}$ is equivalent to applying a monoid action $\varphi(\monoid) \in \monoid^{\prime}$ to the output of the artist $\vartist(\delement) \in \mathscr{\gtotal}$.

@@

\subsubsection{Variables: Fiber Space \dfiber}
\label{sec:data_fiber}
To formalize the structure of the data components, we use notation introduced by Spivak \cite{spivakSIMPLICIALDATABASES} that binds the components of the fiber to variable names and types. Spivak constructs a set \ftotal\ that is the disjoint union of all possible objects of types $\{\ftype_0, \ldots, \ftype_m\} \in \ftypes$, where \ftypes\ are the data types of the variables in the dataset. He then defines the single variable set \fttype\ 
\begin{equation}
    \label{eq:data_types}
    \begin{tikzcd}
        \fttype \arrow[r] \arrow[d, "\pi_{\fsection}"'] & \ftotal \arrow[d, "\pi"] \\
        \fnames \arrow[r, "\fsection"']                          & \ftypes       
    \end{tikzcd}
\end{equation}
which is \ftotal\ restricted to objects of type \ftype\ bound to variable name \fname. The \fttype\ lookup is by name to specify that every component is distinct, since multiple components can have the same type \ftype. Given \fsection, the fiber for a one variable dataset is
\begin{equation}
    \dfiber = \ftotal_{\fsection(\fname)} = \ftotal_{\ftype} 
\end{equation}
where \fsection\ is the schema binding variable name \fname\ to its datatype \ftype. A dataset with multiple variables has a fiber that is the cartesian cross product of $\ftotal_{\fsection}$ applied to all the columns:
\begin{equation}
F = \ftotal_{\fsection(\fname_{1})}\times \ldots \ftotal_{\fsection(\fname_{i})} \ldots\times \ftotal_{\fsection(\fname_{n})}
\end{equation}
which is equivalent to 
\begin{equation}
    \dfiber= \dfiber_{0} \times \ldots \times \dfiber_{i}\times\ldots\times \dfiber_{n}
\end{equation}
which allows us to decouple \dfiber\ into components $\dfiber_i$.

@@

\subsubsection{Continuity: Base Space $K$} 
\label{sec:data_base}
The base space \dbase\ is way to express how the records in \dtotal\ are connected to each other, for example if they are discrete points or if they lie in a 2D continous surface. Connectivity type is assumed in the choice of visualization, for example a line plot implies 1D continuous data, but an explicit representation allows for verifying that the topology of the graphic representation is equivalent to the topology of the data.

@@

\subsubsection{Data: Sections \dsection}
\label{sec:data_section}
The section $\dsection: \dbase\rightarrow \dtotal$ is what ties together the base space \dbase\ with the fiber \dfiber. A section is a function that takes as input location $\dbasepoint \in \dbase$ and returns a record $\delement \in \dtotal$. For example, in the special case of a table \cite{spivakSIMPLICIALDATABASES}, \dbase\ is a set of row ids, \dfiber\ is the columns, and the section \dsection\ returns the record \delement\ at a given key in \dbase. For any fiber bundle, there exists a map
\begin{equation}
    \begin{tikzcd}
        \dfiber \arrow[r, hook] & \dtotal \arrow[d, "\pi"'] \\
                          & \dbase \arrow[u, "\dsection"', bend right]
    \end{tikzcd}
\end{equation}
 such that $\pi(\dsection(\dbasepoint)) = \dbasepoint$. The set of all global sections is denoted as $\Gamma(\dtotal)$. Assuming a trivial fiber bundle $\dtotal = \dbase \times \dfiber$, the section is 
\begin{equation}
    \label{eq:section_return}
    \dsection(\dbasepoint) = (\dbasepoint, (g_{\dfiber_{0}}(\dbasepoint), \ldots, g_{\dfiber_{n}}(\dbasepoint)))
\end{equation}
where $g: \dbase \rightarrow \dfiber$ is the index function into the fiber. This formulation of the section also holds on locally trivial sections of a non-trivial fiber bundle. Because we can decompose the bundle and the fiber, we can decompose \dsection\ as 
\begin{equation}
\label{eq:data_section_decompose}
\dsection= (\dsection_0,\ldots, \dsection_i, \dots, \dsection_n) 
\end{equation}
where each section $\dsection_i$ is a variable or set of variables. This allows for accessing the data component wise rather than just as sections on \dbase.

@@

There are also many types of labeled containers that can richly be described in this framework because of the fiber. For example, a pandas series which stores a labeled list, or a dataframe\cite{jeff_reback_2020_3715232} which stores a relational table. A series could store the values of $\dsection^{(1)}$ and a second series could be  $\dsection^{(2)}$. We could also fatten the fiber to hold two temperature series, such that a section would be an instance of a dataframe with a time column and two temperature columns. While the series and dataframe explicitly have a time index column, they are components in our model and the index is always assumed to be random keys.

@@

\subsection{Graphic: \gtotal}
\label{sec:graphic}  
We introduce a graphic bundle to hold the essential information necessary to render a graphical design constructed by the artist. As with the data, we can represent the target graphic as a section \gsection\ of a bundle  $(\gtotal, \gbase, \pi, \gfiber)$. The graphic bundle \gtotal\ consists of a base \gbase (~\ref{sec:graphic_fiber}) that is a thickened form of \dbase\, 
a fiber \gfiber (~\ref{sec:graphic_base}) that is an idealized display space, and sections \gsection (~\ref{sec:graphic_section}) that encode a graphic where the visual characteristics are fully specified.

@@

\subsubsection{Continuity of the Graphic \gbase} 
\label{sec:graphic_base}
Just as the \dbase\ encodes the connectivity of the records in the data, we propose an equivalent \gbase\ that encodes the connectivity of the rendered elements of the graphic. For some visualizations, \dbase may be lower dimension than \gbase. For example, in a typical 2D display (ignoring depth), a point that is 0D in \dbase cannot be represented on screen unless it is thickened to 2D to encode the connectivity of the pixels that visually represent the point. This thickening is often not necessary when the dimensionality of \dbase\ matches the dimensionality of the target space, for example if \dbase is 2D and the display is a 2D screen. We introduce \gbase\ to thicken \dbase\ in a way which preserves the structure of \dbase.

@@

\subsubsection{Renderering \gsection}
\label{sec:graphic_section}
\begin{figure}[H]
    \includegraphics[width=\textwidth]{figures/math/render.png}
    \caption{To render a graphic, a pixel $p$ is selected in the display space, which is defined in the same coordinates as the x and y components in \gfiber.  The inverse mapping ${\gsection_{xy}}^(p)$ returns a region $\gbase_{p} \subset \gbase$. $\gsection(\gbase_{p})$ returns the list of elements $(x,\,y,\,r,\,g,\,b) \in \gfiber$ that lie over $\gbase_{p}$. The integral over the $(r,\,g,\,b)$ elements is the color of the pixel.}
    \label{fig:graphic_rho_lookup}
\end{figure}

@@

\subsection{Artist}
\label{sec:artist}
We propose that the transformation from data to visual representation can be described as a structure preserving map from one topological space to another. We name this map the artist as that is the analogous part of the  Matplotlib\cite{hunterArchitectureOpenSource} architecture that builds visual elements to pass off to the renderer. The topological artist \vartist\ is a monoid equivariant sheaf map from the sheaf on a data bundle \dtotal\ which is $\mathcal{O}(\dtotal)$ to the sheaf on the graphic bundle \gtotal, $\mathcal{O}(\gtotal)$. 
\begin{equation}
    \vartist: \mathcal{O}(\dtotal) \rightarrow \mathcal{O}(\gtotal)
\end{equation}
Sheafs are a mathematical object with restriction maps that define how to glue \dsection over local neighborhoods $U \subseteq \dbase$,  discussed in section~\ref{sec:artist_sec:data_sheaf_stalk}, such that the \vartist\ maps are consistent over continuous regions of \dbase. While \vartist can usually construct graphical elements solely with the data in \dsection, some visualizations, such as line, may also need some finite number $n$ of derivatives, which is captured by the jet bundle $\mathcal{J}^n$ \cite{JetBundle2020,musilovaCalculusVariationsJet2016} with $\mathcal{J}^{0}(\dtotal)=\dtotal$. In this work, we at most need $\mathcal{J}^{2}(\dtotal)$ which is the value at \dsection\ and its first and second derivatives; therefore the artist takes as input the data bundle padded with the jet bundle $\dtotal^{\prime} = \dtotal + \mathcal{J}^{2}(\dtotal)$.

@@

\subsubsection{Visual Encoders \vchannel}
\label{sec:artist_nu}
%% equivariant maps 
\begin{figure}[H]
    \includegraphics[width=\textwidth]{figures/math/equivariance_nu.png}
    \caption{In this artis, \vchannel\ maps the strings to the emojis. This \vchannel\ is equivariant because the monoid actions (which are represented by the colored arrows) are the same on both the \dsection\ input and \vsection\ output sets.}
    \label{fig:artist_nu}
\end{figure}
As introduced in section~\ref{sec:intro_viz}, there are many ways to visually represent data components. We define the visual transformers \vchannel\ 
\begin{equation}
    \label{eq:nu_expanded}
    \{\vchannel_{0}, \ldots, \vchannel_{n}\}: \{\dsection_{0}, \ldots, \dsection_{n}\} \mapsto \{\vsection_{0}, \ldots, \vsection_{n}\}
\end{equation}
as the set of equivariant maps $\vchannel_i: \dsection_i \mapsto \vsection_i$. Given $\monoid_i$ is the monoid action on $\dtotal_i$ and that there is a monoid ${\monoid_i}^{\prime}$ on $\vtotal_i$, then there is a monoid homomorphism from $\monoid_i \rightarrow {\monoid_i}^{\prime}$ that \vchannel\ must preserve. As mentioned in section~\ref{sec:data_monoid}, we choose monoid actions as the basis for equivariance because they define the structure on the fiber components.

A validly constructed \vchannel\ is one where the diagram of the monoid transform $m$
\begin{equation}
    \label{eq:nu_categorical}
\begin{tikzcd}
    \dtotal_i \arrow[r] \arrow[r, "\vchannel_i"] \arrow[d, "m_{\delement}"'] & \vtotal_i \arrow[d, "m_{\velement}"] \\
    \dtotal_i \arrow[r, "\vchannel_i"]                           & \vtotal_i               
\end{tikzcd}
\end{equation}
commutes such that 
\begin{equation}
\vchannel_i(m_{\delement}(\dtotal_i)) = \varphi(m_{\velement})(\vchannel_i(\dtotal_i))\end{equation}

This equivariance constraint yields guidance on what makes for an invalid transform. For example, given a visual fiber of the same type as the data fiber $\dfiber_i=\vfiber_i$, the transform $\vchannel_{i}(x) = .5$ does not commute under translation $t(x) = x+2$  
\begin{align}
    \vchannel(t(\delement + 2)) & \overset{?}{=} \vchannel(\delement) + \vchannel(2)\\
    .5 &\neq .5 + .5
\end{align}

On the other hand figure~\ref{fig:artist_nu} illustrates a valid \vchannel\ mapping from \textbf{Strings} to symbols. The group action on these sets is permutation, so shuffling the words must have an equivalent shuffle of the symbols they are mapped to. Continuing with the assumption that $\dfiber=\vfiber$, we can describe the constraints on the other Steven's measurment group types. To preserve ordinal and partial order monoid actions, \vchannel\ must be a monotonic function such that given $\delement_1, \delement_2 \in \dtotal_{i}$ 
\begin{equation}
\text{ if } delement_1 \leq \delement_2 \text{ then } \vchannel (\delement_1) \leq \vchannel(\delement_2)
\end{equation}
the visual encodings must also have some sort of ordering. For interval scale data, \vchannel\ is equivariant under translation monoid actions if 
\begin{equation}
\vchannel (x + c) = \vchannel(x) + \vchannel(c)
\end{equation}
while for ratio data, there must be equivalent scaling 
\begin{equation}
\vchannel(xc) = \vchannel(x)\vchannel(c) 
\end{equation}
The assumption $\dfiber_i=\vfiber_i$ is made here so that we could omit  $\varphi: \monoid\rightarrow\monoid^{\prime}$ maps that convert the monoid action on $\dfiber_{i}$ to the equivalent action on $\vfiber_{i}$. As shown in table~\ref{tab:mpl_visual_variable_fiber}, it is not uncommon for data and visual variables to resolve to the same types. The constraints on \vchannel\ can be embedded into our artist such that the \vchannel\ functions are equivariant and also provide guidance on constructing new equivariant \vchannel\ functions.

@@

\subsubsection{Assembly factory \vmarkd}
The graphic base space \gbase\ is not accessible in many architectures, including Matplotlib; instead we can construct a factory function \vmarkd\ over \dbase\ that can build a \vmark. As shown in eq~\ref{eq:artist_diagram}, \vmark\ is a bundle map $\vmark: \vtotalpull\rightarrow \gtotal$ where \vtotalpull\ and \gtotal\ are both bundles over \gbase.
\begin{figure}[H]
    \includegraphics[width=1\textwidth]{figures/math/q_hat.png}
    \caption{The pullback of the visual bundle \vtotalpull\ is the replication of a \vsection\ over all points \gbasepoint\ that map back to a single \dbasepoint. Because the \vsection\ is the same, we can constract a \vmarkd\ on \vsection\ over \dbasepoint\ that will fabricate the \vmark\ for the equivalent set of \gbasepoint\ associated to that \dbasepoint}
    \label{fig:artist_q_hat}
\end{figure}

@@

\subsubsection{Sheaf}
\label{sec:artist_sheaf_stalk}
The restriction maps of a sheaf describe how local \dsection can be glued into larger sections \cite{ghristElementaryAppliedTopology2014,ghristHomologicalAlgebraData2018}. As part of the definition of local triviality, there is an open neighborhood $U\subset \dbase$ for every $\dbasepoint \in \dbase$. We can define the inclusion map $\iota: U \rightarrow \dbase$  which pulls \dtotal over $U$ 
\begin{equation}
    \label{eq:sheaf}
    \begin{tikzcd}
        \iota^*\dtotal \arrow[d, "\pi"'] \arrow[r, "\iota^*", hook]             & \dtotal \arrow[d, "\pi"']                  \\
        U \arrow[r, "\iota", hook] \arrow[u, "\iota^{*}\dsection"', bend right] & \dbase \arrow[u, "\dsection"', bend right]
    \end{tikzcd}
\end{equation}
such that the pulled back $\iota^*\dsection$ only contains records over $U \subset \dbase$. By gluing $\iota^*\dsection$ together, the sheaf is putting a continuous structure on local sections which allows for defining a section over a subset in \dbase. That section over subset \dbase\ maps to the graphic generated by \vartist\ for visualizations such as sliding windows\cite{crouchDynamicGraphsSlidingwindow2013,chuTimeSeriesSegmentation1995} streaming data, or navigation techniques such as pan and zoom\cite{NekrasovskiEvaluationPanZoom2006}.

>>>>

As discussed in the introduction, visualization is generally defined as structure preserving maps from a data object to a graphic object. In order to formalize this statement, we describe the connectivity of the records using topology and define the structure on the components in terms of the monoid actions on the component types. By formalizing structure in this way, we can evaluate the extent to which a visualization preserves the structure of the data it is representing and build structure preserving visualization tools. 
We introduce the notion of an artist $\mathscr{\vartist}$ as an equivariant map from data to graphic
\begin{equation}
    \label{eq:artist}
    \mathscr{\vartist}: \mathscr{\dtotal} \rightarrow \mathscr{\gtotal}
\end{equation}
that carries a homomorphism of monoid actions $\varphi: \monoid \rightarrow \monoid^{\prime}$ \cite{cegarraCohomologyMonoidsOperators2019}, which are discussed in detail in section~\ref{sec:data_monoid}. Given \monoid\ on data $\mathscr{\dtotal}$ and $\monoid^{\prime}$ on graphic $\mathscr{\gtotal}$, we propose that artists $\mathscr{\vartist}$ are equivariant maps 
\begin{equation}
\mathscr{\vartist}(m\cdot \delement) = \varphi(m)\cdot\mathscr{\vartist}(\delement) 
\end{equation}
such that applying a monoid action $m \in \monoid$ to the data $\delement \in \mathscr{\dtotal}$ input to $\mathscr{\vartist}$ is equivalent to applying a monoid action $\varphi(\monoid) \in \monoid^{\prime}$ to the graphic $\vartist(\delement) \in \mathscr{\gtotal}$ output of the artist.

@@

\subsubsection{Variables in Fiber Space \dfiber}
\label{sec:data_fiber}
To formalize the structure of the data components, we use notation introduced by Spivak \cite{spivakSIMPLICIALDATABASES} that binds the components of the fiber to variable names and types. Spivak constructs a set \ftotal\ that is the disjoint union of all possible objects of types $\{\ftype_0, \ldots, \ftype_m\} \in \ftypes$, where \ftypes\ are the data types of the variables in the dataset. He then defines the single variable set \fttype\ 
\begin{equation}
    \label{eq:data_types}
    \begin{tikzcd}
        \fttype \arrow[r] \arrow[d, "\pi_{\fsection}"'] & \ftotal \arrow[d, "\pi"] \\
        \fnames \arrow[r, "\fsection"']                          & \ftypes       
    \end{tikzcd}
\end{equation}
which is \ftotal\ restricted to objects of type \ftype\ bound to variable name \fname. The \fttype\ lookup is by name to specify that every component is distinct, since multiple components can have the same type \ftype. Given \fsection, the fiber for a one variable dataset is
\begin{equation}
    \dfiber = \ftotal_{\fsection(\fname)} = \ftotal_{\ftype} 
\end{equation}
where \fsection\ is the schema binding variable name \fname\ to its datatype \ftype. A dataset with multiple variables has a fiber that is the cartesian cross product of $\ftotal_{\fsection}$ applied to all the columns:
\begin{equation}
F = \ftotal_{\fsection(\fname_{1})}\times \ldots \ftotal_{\fsection(\fname_{i})} \ldots\times \ftotal_{\fsection(\fname_{n})}
\end{equation}
which is equivalent to 
\begin{equation}
    \dfiber= \dfiber_{0} \times \ldots \times \dfiber_{i}\times\ldots\times \dfiber_{n}
\end{equation}
which allows us to decouple \dfiber\ into components $\dfiber_i$.

@@

\subsubsection{Continuity of the Data $K$} 
\label{sec:data_base}
The base space \dbase\ is way to express how the records in \dtotal\ are connected to each other, for example if they are discrete points or if they lie in a 2D continous surface. Connectivity type is assumed in the choice of visualization, for example a line plot implies 1D continuous data, but an explicit representation allows for verifying that the topology of the graphic representation is equivalent to the topology of the data.

@@

\subsubsection{Data \dsection}
\label{sec:data_section}
While the projection function $\pi:\dtotal \rightarrow\dbase$ ties together the base space \dbase\ with the fiber \dfiber, a section $\dsection: \dbase\rightarrow \dtotal$ encodes a dataset. A section function takes as input location $\dbasepoint \in \dbase$ and returns a record $\delement \in \dtotal$. For example, in the special case of a table \cite{spivakSIMPLICIALDATABASES}, \dbase\ is a set of row ids, \dfiber\ is the columns, and the section \dsection\ returns the record \delement\ at a given key in \dbase. For any fiber bundle, there exists a map
\begin{equation}
    \begin{tikzcd}
        \dfiber \arrow[r, hook] & \dtotal \arrow[d, "\pi"'] \\
                          & \dbase \arrow[u, "\dsection"', bend right]
    \end{tikzcd}
\end{equation}
 such that $\pi(\dsection(\dbasepoint)) = \dbasepoint$. The set of all global sections is denoted as $\Gamma(\dtotal)$. Assuming a trivial fiber bundle $\dtotal = \dbase \times \dfiber$, the section is 
\begin{equation}
    \label{eq:section_return}
    \dsection(\dbasepoint) = (\dbasepoint, (g_{\dfiber_{0}}(\dbasepoint), \ldots, g_{\dfiber_{n}}(\dbasepoint)))
\end{equation}
where $g: \dbase \rightarrow \dfiber$ is the index function into the fiber. This formulation of the section also holds on locally trivial sections of a non-trivial fiber bundle. Because we can decompose the bundle and the fiber, we can decompose \dsection\ as 
\begin{equation}
\label{eq:data_section_decompose}
\dsection= (\dsection_0,\ldots, \dsection_i, \dots, \dsection_n) 
\end{equation}
where each section $\dsection_i$ is a variable or set of variables. This allows for accessing the data component wise rather than just as sections on \dbase.

@@

There are also many types of labeled containers that can richly be described in this framework because of the fiber. For example, a pandas series which stores a labeled list, or a dataframe\cite{jeff_reback_2020_3715232} which stores a relational table. A series could store the values of $\dsection^{(1)}$ and a second series could be  $\dsection^{(2)}$. We could also fatten the fiber to hold two temperature series, such that a section would be an instance of a dataframe with a time column and two temperature columns. While the series and dataframe explicitly have a time index column, they are components in our model and the index is assumed to be data independent references such as hashvalues, virtual memory locations, or random number keys.

@@

\subsection{Graphic Space \gtotal}
\label{sec:graphic}  
We introduce a graphic bundle to hold the essential information necessary to render a graphical design constructed by the artist. As with the data, we can represent the target graphic as a section \gsection\ of a bundle  $(\gtotal, \gbase, \pi, \gfiber)$. The graphic bundle \gtotal\ consists of a base \gbase (~\ref{sec:graphic_fiber}) that is a thickened form of \dbase\, 
a fiber \gfiber (~\ref{sec:graphic_base}) that is an idealized display space, and sections \gsection (~\ref{sec:graphic_section}) that encode a graphic where the visual characteristics are fully specified.

@@

\subsubsection{Continuity of the Graphic \gbase} 
\label{sec:graphic_base}
Just as the \dbase\ encodes the connectivity of the records in the data, we propose an equivalent \gbase\ that encodes the connectivity of the rendered elements of the graphic. For example, consider a \gbase\ that is mapped to the region of a 2D display space that represents \dbase. For some visualizations, \dbase\ may be lower dimension than \gbase. For example, a point that is 0D in \dbase\ cannot be represented on screen unless it is thickened to 2D to encode the connectivity of the pixels that visually represent the point. This thickening is often not necessary when the dimensionality of \dbase\ matches the dimensionality of the target space, for example if \dbase\ is 2D and the display is a 2D screen. We introduce \gbase\ to thicken \dbase\ in a way which preserves the structure of \dbase.

@@

\subsubsection{Graphic \gsection}
\label{sec:graphic_section}
\begin{figure}[H]
    \includegraphics[width=\textwidth]{figures/math/render.png}
    \caption{To render a graphic, a pixel $p$ is selected in the display space, which is defined in the same coordinates as the x and y components in \gfiber.  The inverse mapping ${\gsection_{xy}}^(p)$ returns a region $\gbase_{p} \subset \gbase$. $\gsection(\gbase_{p})$ returns the list of elements $(x,\,y,\,r,\,g,\,b) \in \gfiber$ that lie over $\gbase_{p}$. The integral over the $(r,\,g,\,b)$ elements is the color of the pixel.}
    \label{fig:graphic_rho_lookup}
\end{figure}

@@

\subsection{Artist}
\label{sec:artist}
We propose that the transformation from data to visual representation can be described as a structure preserving map from one topological space to another. We name this map the artist as that is the analogous part of the  Matplotlib\cite{hunterArchitectureOpenSource} architecture that builds visual elements to pass off to the renderer. The topological artist \vartist\ is a monoid equivariant sheaf map from the sheaf on a data bundle \dtotal\ which is $\mathcal{O}(\dtotal)$ to the sheaf on the graphic bundle \gtotal, $\mathcal{O}(\gtotal)$. 
\begin{equation}
    \vartist: \mathcal{O}(\dtotal) \rightarrow \mathcal{O}(\gtotal)
\end{equation}
Sheafs are a mathematical object with restriction maps that define how to glue \dsection over local neighborhoods $U \subseteq \dbase$,  discussed in section~\ref{sec:artist_sec:data_sheaf_stalk}, such that the \vartist\ maps are consistent over continuous regions of \dbase. While \vartist can usually construct graphical elements solely with the data in \dsection, some visualizations, such as line, may also need some finite number $n$ of derivatives, which is captured by the jet bundle $\mathcal{J}^n$ \cite{JetBundle2020,musilovaCalculusVariationsJet2016} with $\mathcal{J}^{0}(\dtotal)=\dtotal$. In this work, we at most need $\mathcal{J}^{2}(\dtotal)$ which is the value at \dsection\ and its first and second derivatives; therefore the artist takes as input the jet bundle $\dotal^{\prime}=\mathcal{J}^{2}(\dtotal)$.

@@

\subsubsection{Visual Encoders \vchannel}
\label{sec:artist_nu}
%% equivariant maps 
\begin{figure}[H]
    \includegraphics[width=\textwidth]{figures/math/equivariance_nu.png}
    \caption{In this artis, \vchannel\ maps the strings to the emojis. This \vchannel\ is equivariant because the monoid actions (which are represented by the colored arrows) are the same on both the \dsection\ input and \vsection\ output sets.}
    \label{fig:artist_nu}
\end{figure}
As introduced in section~\ref{sec:intro_viz}, there are many ways to visually represent data components. We define the visual transformers \vchannel\ 
\begin{equation}
    \label{eq:nu_expanded}
    \{\vchannel_{0}, \ldots, \vchannel_{n}\}: \{\dsection_{0}, \ldots, \dsection_{n}\} \mapsto \{\vsection_{0}, \ldots, \vsection_{n}\}
\end{equation}
as the set of equivariant maps $\vchannel_i: \dsection_i \mapsto \vsection_i$. Given $\monoid_i$ is the monoid action on $\dtotal_i$ and that there is a monoid ${\monoid_i}^{\prime}$ on $\vtotal_i$, then there is a monoid homomorphism from $\varphi:\monoid_i \rightarrow {\monoid_i}^{\prime}$ that \vchannel\ must preserve. As mentioned in section~\ref{sec:data_monoid}, we choose monoid actions as the basis for equivariance because they define the structure on the fiber components.

A validly constructed \vchannel\ is one where the diagram of the monoid transform $m$ commutes such that 
\begin{equation}
    \label{eq:nu_categorical}
\begin{tikzcd}
    \dtotal_i \arrow[r] \arrow[r, "\vchannel_i"] \arrow[d, "m_{\delement}"'] & \vtotal_i \arrow[d, "m_{\velement}"] \\
    \dtotal_i \arrow[r, "\vchannel_i"]                           & \vtotal_i               
\end{tikzcd}
\end{equation}

In general, the data fiber $\dfiber_{i}$ cannot be assumed to be of the same type as the visual fiber $\vfiber_{i}$; therefore the actions of \monoid\ on  $\dfiber_{i}$ are not the same as the actions of $\moniod^{prime}$ on \vfiber.  This means that an equivariant \vchannel\ must commute
\begin{equation}
\vchannel_i(m_{\delement}(\dtotal_i)) = \varphi(m_{\delement})(\vchannel_i(\dtotal_i))\end{equation} 
such that $\varphi: m_{\delement} \mapsto m_{\velements}$ maps a monoid action on data to a monoid action on visual elements. However, we can construct a monoid action of \monoid\ on $\vfiber_{i}$ that is compatible with a monoid action of \monoid\ on $\dfiber_{i}$. We can compose the monoid actions on the visual fiber $\monoid^{\prime} \times \vfiber_{i} \rightarrow \vfiber_{i}$ with the homomorphism $\varphi: \monoid \rightarrow \monoid^{\prime}$ that takes \monoid\ to $\monoid^{\prime}$. This allows us to define a monoid action on \vfiber\ of \monoid\ that is $(m, \velement) \rightarrow \varphi(m)\bullet\velement$. Therefore without a loss of generality, we can assume that \monoid\ acts on $\dfiber_{i}$ and on $\vfiber_{i}$ compatibly such that $\varphi$ is the identity function.

On example of an equivariant \vchannel\ is illustrated in figure~\ref{fig:artist_nu}, which is a mapping from \textbf{Strings} to symbols. This is also an example of a Steven's nominal measurement set, which is defined as having permutation group actions 
\begin{equation}
    \text{ if } \delement_1 \neq \delement_2 \text{ then } \vchannel (\delement_1) \neq\vchannel(\delement_2)
\end{equation}
such that shuffling the words must have an equivalent shuffle of the symbols they are mapped to. This is illustrated in the identical actions, represented by the colored arrows, on the words and emojis. To preserve ordinal and partial order monoid actions, \vchannel\ must be a monotonic function such that given $\delement_1, \delement_2 \in \dtotal_{i}$,
\begin{equation}
\text{ if } \delement_1 \leq \delement_2 \text{ then } \vchannel (\delement_1) \leq \vchannel(\delement_2)
\end{equation}
the visual encodings must also have some sort of ordering. For interval scale data, \vchannel\ is equivariant under translation monoid actions if 
\begin{equation}
\vchannel (x + c) = \vchannel(x) + c
\end{equation}
while for ratio data, there must be equivalent scaling 
\begin{equation}
\vchannel(xc) = \vchannel(x)*c 
\end{equation}

We therefore can test if a \vchannel\ is equivariant by testing the actions under which is must  commute. For example, we define a transform $\vchannel_{i}(x) = .5$ on interval data. This means it must commute under translation, for example $t(x) = x+2$. Testing this constraint 
\begin{align}
    \vchannel(t(\delement + 2)) & \overset{?}{=} \vchannel(\delement) + \varphi(2)\\
    .5 &\neq .5 + 2
\end{align}
we find that the \vchannel\ defined here does not commute and is therefore invalid. The constraints on \vchannel\ can be embedded into our artist such that the \vchannel\ functions are equivariant and also provide guidance on constructing new equivariant \vchannel\ functions.

@@

\subsubsection{Assembly factory \vmarkd}
The graphic base space \gbase\ is not accessible in many architectures, including Matplotlib; instead we can construct a factory function \vmarkd\ over \dbase\ that can build a \vmark. As shown in eq~\ref{eq:artist_diagram}, \vmark\ is a bundle map $\vmark: \vtotalpull\rightarrow \gtotal$ where \vtotalpull\ and \gtotal\ are both bundles over \gbase.
\begin{figure}[H]
    \includegraphics[width=1\textwidth]{figures/math/q_hat.png}
    \caption{The pullback of the visual bundle \vtotalpull\ is the replication of a \vsection\ over all points \gbasepoint\ that map back to a single \dbasepoint. Because the \vsection\ is the same, we can constract a \vmarkd\ on \vsection\ over \dbasepoint\ that will fabricate the \vmark\ for the equivalent region of \gbasepoint\ associated to that \dbasepoint}
    \label{fig:artist_q_hat}
\end{figure}

@@

\subsubsection{Sheafs}
\label{sec:artist_sheaf_stalk}
The restriction maps of a sheaf describe how local \dsection can be glued into larger sections \cite{ghristElementaryAppliedTopology2014,ghristHomologicalAlgebraData2018}. As part of the definition of local triviality, there is an open neighborhood $U\subset \dbase$ for every $\dbasepoint \in \dbase$. We can define the inclusion map $\iota: U \rightarrow \dbase$  which pulls \dtotal over $U$ 
\begin{equation}
    \label{eq:sheaf}
    \begin{tikzcd}
        \iota^*\dtotal \arrow[d, "\pi"'] \arrow[r, "\iota^*", hook]             & \dtotal \arrow[d, "\pi"']                  \\
        U \arrow[r, "\iota", hook] \arrow[u, "\iota^{*}\dsection"', bend right] & \dbase \arrow[u, "\dsection"', bend right]
    \end{tikzcd}
\end{equation}
such that the pulled back $\iota^*\dsection$ only contains records over $U \subset \dbase$. By gluing $\iota^*\dsection$ together, the sheaf is putting a continuous structure on local sections which allows for defining a section over a subset in \dbase. That section over subset \dbase\ maps to the graphic generated by \vartist\ for visualizations such as sliding windows\cite{crouchDynamicGraphsSlidingwindow2013,chuTimeSeriesSegmentation1995} streaming data, or navigation techniques such as pan and zoom\cite{NekrasovskiEvaluationPanZoom2006}.
